### 2015-01-28 comparaison.tex
update intermediate version
update intermediate version

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al. \cite{roccaasp} which was developed also on ASP but for transition states networks.
Each method use a specific kind of representation\footnote{The transalation from PH to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transitions network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & out & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Diviser en 2 tableaux}

\todo{Il n'y a pas de résultats sur le calcul des poitns fixes dans le tableau} \repEB{En fait pour la méthode de Rocca le script qu'il m'ont envoyé génére quelques erreurs lors de la compilation. pour libddd y'avait pas l'option (--fixpoint) ne génère rien comme résultat et GINsim répond très rapidement mais il faudrait calculer tout le graphe de transition avant. }

@@

\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
of state-transition networks.
We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
Therefore, using this method first requires an exponential computation to
express the state-transition graph of the studied model into ASP.
Indeed, for the tadpole tail resorption model comprising only 42 processes,
this computation took 3 minutes, which was not included in the solving time
reported in table~\ref{tab:reachability}.
Afterwards, this method is able to check any kind of CTL formula
(and not only the EF form that we focused on in this paper),
but this checking may be time-consuming.

@@

In the previous sections,
we developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
For example, the total number of states is an obvious maximum,
as it will never be exceeded by a minimum path,
but it is too hight to be very interesting.
The total number of sorts is a more interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks;
or, with a similar reasoning, the total number of processes can be chosen.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
\todo{Y a-t-il moyen de les différencier de chemins corrects?}

>>>>

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al. \cite{roccaasp} which was developed also on ASP but for transition states networks.
Each method use a specific kind of representation\footnote{The transalation from PH from or to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transitions network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & \todo{XX} & Inconc & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & \todo{XX}  & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & 0m2.61s & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & Inconc & out & out & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Diviser en 2 tableaux ou c'est bon??}

@@

\item \textbf{ASP-CTL} \todo{à revoir} offers the possibility to model-check CTL properties
of state-transition networks.
We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
Therefore, using this method first requires an exponential computation to
express the state-transition graph of the studied model into ASP.
Indeed, for the tadpole tail resorption model comprising only 42 processes,
this computation took 3 minutes, which was not included in the solving time
reported in table~\ref{tab:reachability}.
Afterwards, this method is able to check any kind of CTL formula
(and not only the EF form that we focused on in this paper),
but this checking may be time-consuming.

@@

In the previous sections,
we developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option \texttt{"-{}-imax=n"} of \textsc{iclingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
For example, the total number of states is an obvious maximum,
as it will never be exceeded by a minimum path,
but it is too hight to be very interesting.
The total number of sorts is a more interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks;
or, with a similar reasoning, the total number of processes can be chosen.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and there is no path to be displayed because there is no path verifies the goals after \texttt{n} changes.

### 2015-01-28 fixPoint.tex
update intermediate version
update intermediate version

\subsection{Process Hitting translation in ASP}
Before writing any piece of scripts on ASP to analyse the dynamic of the network,
we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods, translation from PH to ASP and the dynamic verification on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.

@@

\begin{example}[Representation of a PH network in ASP]
\label{ex:asp-ph}
The representation of the PH network of figure \ref{fig:ph} in ASP is the following:
\begin{lstlisting}
sort(X) :- process(X,) %\label{ASPsort}
process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
action("b",0,"z",1,2). action("z",0,"a",0,1). %\label{actions2}
\end{lstlisting}
Line \ref{ASPsort} enumerates every sort of network.
In line \ref{ASPprocess} we create the list of processes corresponding to each sort,
for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2};
this specific predicate will in fact expand into the three following predicates:
\texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
for example, the first predicate \texttt{action("a",0,"b",1,0)} represents the action
$\PHfrappe{a_1}{b_1}{b_0}$.
\end{example}

\subsection{Search of fixed points}
\todo{ Parler du point fixe et ses interrêts. \\
Décrire la nouvelle méthode qui permet de déterminer les états stables d'un réseau }

>>>>

\subsection{Process Hitting translation in ASP}
Before writing any piece of scripts on ASP to analyse the dynamic of the network,
we first need to translat the concerned PH network in ASP  \footnote{You can find our new methods for the dynamic verification and also the translation from PH to ASP on line: \url{https://github.com/EmnaBenAbdallah/verification-of-dynamical-properties_PH}}.

@@

\begin{example}[Representation of a PH network in ASP]
\label{ex:asp-ph}
The representation of the PH network of figure \ref{fig:ph} in ASP is the following:
\begin{lstlisting}
sort(X) :- process(X,I) %\label{ASPsort}
process("a", 0..1). process("b", 0..1). process("z", 0..2). %\label{ASPprocess}
action("a",0,"b",1,0). action("a",1,"a",1,0). action("b",1,"z",0,2). %\label{actions1}
action("b",0,"z",1,2). action("z",0,"a",0,1). %\label{actions2}
\end{lstlisting}
Line \ref{ASPsort} enumerates every sort of network.
In line \ref{ASPprocess} we create the list of processes corresponding to each sort,
for example the sort \texttt{"z"} has 3 processes numbered from \texttt{0} to \texttt{2};
this specific predicate will in fact expand into the three following predicates:
\texttt{process("z", 0)}, \texttt{process("z", 1)}, \texttt{process("z", 2)}.
Finally, all the actions of the network are defined in lines \ref{actions1} and \ref{actions2};
for example, the first predicate \texttt{action("a",0,"b",1,0)} represents the action
$\PHfrappe{a_1}{b_1}{b_0}$.
\end{example}

\subsection{Search of fixed points}
